---
title: "An Introduction to Text Analysis"
author: "Ayush Patel"
subtitle: "DDD: Elements of Statistical Machine Learning & Politics of Data"
institute: "At Azim Premji University"
date: today
date-format: "DD MMM, YYYY"
format: 
  revealjs:
    fig-height: 6
    fig-width: 14
    embed-resources: true
    margin-left: 50px
    margin-right: 50px
    slide-number: c/t
    width: 1400
    height: 850
    theme: [default, theme.scss]
    footer: "Email: ayush.ap58@gmail.com"
---

```{r}
#| include: false
#| warning: false

library(tidyverse)
library(tidytext)
```

# Hello

::::{.columns}

:::{.column}

[I am Ayush.]{.fragment fragment-index="1" style="font-size:45px"}

[I am a researcher working at the intersection of data, development and economics.]{.fragment fragment-index="2" style="font-size:25px"}

[I am a [RStudio (Posit) certified tidyverse Instructor.](https://education.rstudio.com/trainers/people/patel+ayush/)]{.fragment fragment-index="3" style="font-size:25px"}

[I am a Researcher at [Oxford Poverty and Human development Initiative (OPHI)](https://ophi.org.uk/), at the University of Oxford.]{.fragment fragment-index="4" style="font-size:25px"}

:::

:::{.column}

![[Hello (Eh bonjour– donc–.) by Charles Motte](https://www.metmuseum.org/art/collection/search/392118"Well)](https://images.metmuseum.org/CRDImages/dp/original/DP808141.jpg){fig-align="center" height=400 .lightbox}
 

:::

::::

# Did you come prepared?

::::{.columns}

:::{.column}
- You have installed R. If not see [this link]().

- You have installed RStudio/Positron/VScode or any other IDE. It is recommended that you work through an IDE

- You have the libraries `{tidytext} ` installed

:::

:::{.column}

![[Armor (Gusoku) Helmet signed by Bamen Tomotsugu](https://www.metmuseum.org/art/collection/search/24975)](https://images.metmuseum.org/CRDImages/aa/original/DT5333.jpg){fig-align="center" height=400 .lightbox}


:::

::::


# Before NLP and Text Analysis

[One needs to know how to mine, warngle, and manipulate text data.]{.fragment fragment-index='1'}
<br><br>
[tidy priciples can make this task easy (relatively)]{.fragment fragment-index='2'}
<br><br>
[*"Treating text as data frames of individual words allows us to manipulate, summarize, and visualize the characteristics of text easily and integrate natural language processing into effective workflows..."*]{.fragment fragment-index='3'}
<br><br>
[We do this using the {tidytext} alonmg with other know methods]{.fragment fragment-index='4'}

# Tidy Data

As described in [Wickham 2014](https://www.jstatsoft.org/article/view/v059i10)

1. Each variable is a column.
2. Each observation is a row.
3. Each type of observational unit is a table.

# Tidy Text

As described by [Silge and Robinson 2017](https://www.tidytextmining.com/)

1. **A table with one-token-per-row**
2. token: A meaningful unit of text. (word,sentence,etc)
3. There are other forms of text data structures: string, corpus, document-term matric. (more on this later)

# Tokens

>"*A token is a meaningful unit of text, most often a word, that we are interested in using for further analysis, and tokenization is the process of splitting text into tokens.*"

# Unnesting Tokens

We need to get the text in *one-token-per-row* to acheive the Tidy Text format. This can be dome by the `unnest_tokens()` function from {tidytext}
<br>

```{r}
#| echo: true

text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")

text

```

# Unnesting Tokens

Notice any punctuations?
<br>
```{r}
tibble(
  row = c(1:length(text)),
  text = text
) |>
  unnest_tokens(token = "words",output = word, input =  text)
```

# The RBI MPC minutes


